\chapter*{Acknowledgment}
First of all, I want to thank my advisor Prof. Dr. Jens Niemeyer for offering
me the opportunity to join the 
International Research Training Group "Theoretical Astrophysics and Particle
Physics" and to work on the forefront of turbulence research in astrophysics.
His supervision and amazing intuition for physics always brought me back on
track when my work came to a standstill.

I'm indebted to Dr. Wolfram Schmidt, for patiently answering a
plethora of questions about his subgrid model and turbulence modeling in
general. This work would have been impossible without his contributions.
Also many thanks go to Dr. Luigi Iapichino for introducing me into the physics
of galaxy clusters. Furthermore he had the astonishing talent of always finding
the most interesting literature related to my work.

Many people contributed to the cosmological fluid code Enzo, which I used for
this work: Dr. Wolfram Schmidt integrated the routines for random forcing;
without them, driven turbulence simulation wouldn't been possible.
Christoph Federrath implemented most of the routines for the statistics tool,
which I heavily used for this work. Significant bug fixes and code contributions
came from Dr. Luigi Iapichino and Markus Hupp. Also, important help was 
provided by the original developers of Enzo from the other side of the ocean,
namely Alexei Kritsuk, Brian O'Shea, Robert Harkness, and David Collins. 
I thank all these people for their support. 

Special thanks go to Greg Harper for proof-reading my thesis. I greatly
appreciate his help. 

I'm very much obliged to my parents, who patiently supported me
during the entire duration of my study. But most of all I have to thank my
girlfriend Olga, who never stopped encouraging me and sweetened this
stressful time of life. 
 




  